### 2017-12-21 README-norefs.Rmd

---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`methyvim`

[![Travis-CI Build Status](https://travis-ci.org/nhejazi/methyvim.svg?branch=master)](https://travis-ci.org/nhejazi/methyvim)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/nhejazi/methyvim?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/methyvim)
[![Coverage Status](https://img.shields.io/codecov/c/github/nhejazi/methyvim/master.svg)](https://codecov.io/github/nhejazi/methyvim?branch=master)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)
[![DOI](https://zenodo.org/badge/79256902.svg)](https://zenodo.org/badge/latestdoi/79256902)

> Targeted Data-Adaptive Estimation and Inference for Differential Methylation
> Analysis

__Author:__ [Nima Hejazi](http://nimahejazi.org)

---

## What's `methyvim`?

`methyvim` is an R package that provides facilities for differential methylation
analysis based on _variable importance measures_ (VIMs), a class of
statistically estimable target parameters that arise in causal inference.

The statistical methodology implemented computes targeted minimum loss-based
estimates of several well-characterized variable importance measures:

For discrete-valued treatments or exposures:

* The _average treatment effect_ (ATE): The effect of a binary exposure or
  treatment on the observed methylation at a target CpG site is estimated,
  controlling for the observed methylation at all other CpG sites in the same
  neighborhood as the target site, based on an additive form. In particular, the
  parameter estimate represents the __additive difference__ in methylation that
  would have been observed at the target site had all observations received the
  treatment versus the scenario in which none received the treatment.

* The _relative risk_ (RR): The effect of a binary exposure or treatment on the
  observed methylation at a target CpG site is estimated, controlling for the
  observed methylation at all other CpG sites in the same neighborhood as the
  target site, based on an geometric form. In particular, the parameter estimate
  represents the __multiplicative difference__ in methylation that would have
  been observed at the target site had all observations received the treatment
  versus the scenario in which none received the treatment.

For continous-valued treatments or exposures:

* A _nonparametric variable importance measure_ (NPVI):
  The effect of continous-valued exposure or treatment (the observed methylation
  at a target CpG site) on an outcome of interest is estimated, controlling for
  the observed methylation at all other CpG sites in the same neighborhood as
  the target (treatment) site, based on a parameter that compares values of the
  treatment against a reference value taken to be the null. In particular, the
  implementation provided is designed to assess the effect of differential
  methylation at the target CpG site on a (typically) phenotype-level outcome
  of interest (e.g., survival), in effect providing an nonparametric evaluation
  of the impact of methylation at the target site on said outcome.

_In all cases, an estimator of the target parameter is constructed via targeted
minimum loss-based estimation._

These methods allow differential methylation effects to be quantified in a
manner that is largely free of assumptions, especially of the variety exploited
in parametric models. __The statistical algorithm consists in several major
steps:__

1. Pre-screening of genomic sites is used to isolate a subset of sites for
which there is cursory evidence of differential methylation. For the sake of
computational feasibility, targeted minimum loss-based estimates of VIMs are
computed only for this subset of sites. Several screening approaches are
available, adapting core routines from the following R packages:
[`limma`](http://bioconductor.org/packages/release/bioc/html/limma.html),
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi).
2. Nonparametric VIMs are estimated for the specified parameter, currently
adapting routines from the
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi) and
[`tmle`](https://CRAN.R-project.org/package=tmle) R packages.
3. Since pre-screening is performed prior to estimating VIMs, we make use of a
multiple testing correction uniquely suited to such settings. Due to the
multiple testing nature of the estimation problem, a variant of the Benjamini &
Hochberg procedure for controlling the False Discovery Rate (FDR) is applied.
Specifically, we apply the modified marginal Benjamini & Hochberg step-up False
Discovery Rate controlling procedure for multi-stage analyses (FDR-MSA).

---

## Installation

For standard use, install from [Bioconductor](https://bioconductor.org):

```{r bioc-installation, eval = FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite("methyvim")
```

To contribute, install the bleeding-edge _development version_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-master-installation, eval = FALSE}
devtools::install_github("nhejazi/methyvim")
```

Current and prior [Bioconductor](https://bioconductor.org) releases are
available under branches with numbers prefixed by "RELEASE_". For example, to
install the version of this package available via Bioconductor 3.6, use

```{r gh-develop-installation, eval = FALSE}
devtools::install_github("nhejazi/methyvim", ref = "RELEASE_3_6")
```

---

<!--
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```
-->

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/nhejazi/methyvim/issues).

---

## Contributions

It is our hope that `methyvim` will grow to be widely adopted as a tool for the
nonparametric assessment of variable importance in studies of differential
methylation. To that end, contributions are very welcome, though we ask that
interested contributors consult our [`contribution
guidelines`](https://github.com/nhejazi/methyvim/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `methyvim` R package, please cite the following:

        @article{hejazi2017methyvim,
          doi = {},
          url = {},
          year  = {2017},
          month = {},
          publisher = {},
          volume = {},
          author = {Hejazi, Nima S and Hubbard, Alan E and {van der Laan}, Mark
            J},
          title = {methyvim: Targeted and model-free differential methylation
            analysis in R},
          journal = {}
        }

---

## Related

* [R/`methyvimData`](https://github.com/nhejazi/methyvimData) - R package with
    sample experimental DNA methylation data for use as an example with this
    analysis package.

---

## Funding

The development of this software was supported in part through a grant from the
National Library of Medicine of the NIH (T32 LM012417).

---

## License

&copy; 2017 [Nima S. Hejazi](http://nimahejazi.org)

The contents of this repository are distributed under the MIT license. See file
`LICENSE` for details.