### 2018-04-16 paper/paper_BiocF1000.tex
F1000 (#34)
* use minfi data

* add paper refs

* started to add a bit to the introduction &  moved the text from the repository to the appropriate sections in the article. I will use the text from the repo as a rough guide to draft something more cohesive

* revisions/updates to F1000 manuscript (#28)

* minor vignette additions/corrections

* almost done with f1000, minor vignette corrections

* adding the Summary and final touches

* a few more minor things

* fixing merge conflict

* minor formatting

* suggestion: convert betas to mvals in the limma screenas the mvals have far more centra

\documentclass[9pt,a4paper,]{extarticle}

\usepackage{f1000_styles}

\usepackage[pdfborder={0 0 0}]{hyperref}

\usepackage[numbers]{natbib}
\bibliographystyle{unsrtnat}

%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% disable code chunks background
%\renewenvironment{Shaded}{}{}

% disable section numbers
\setcounter{secnumdepth}{0}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\pagestyle{front}

\title{\texttt{methyvim}: Targeted and Model-free Differential Methylation Analysis in R}

\author[1,2]{Nima S. Hejazi\thanks{\ttfamily nhejazi@berkeley.edu}}
\author[1]{Rachael V. Phillips}
\author[1]{Alan E. Hubbard}
\author[1,3]{Mark J. van der Laan}
\affil[1]{Group in Biostatistics, University of California, Berkeley}
\affil[2]{Center for Computational Biology, University of California, Berkeley}
\affil[3]{Department of Statistics, University of California, Berkeley}

\maketitle
\thispagestyle{front}

\begin{abstract}
We present \texttt{methyvim}, a general algorithm for the nonparametric estimation of
treatment/exposure effects on DNA methylation at CpG sites, complete with
straightforward statistical inference for such estimates. This approach
leverages variable importance measures, a class of parameters that arise in
causal inference and are defined in such a manner that they provide targeted
estimates of the relative importance of individual CpG site with respect to
a binary exposure/treatment assigned at the level of subjects. Additionally,
this procedure is computationally efficient; incorporating a preliminary
screening step to isolate a subset of sites for which there is cursory
evidence of differential methylation and a multiple testing correction step
to control the False Discovery Rate as if all sites were tested. This
technique for analysis of differentially methylated positions provides an
avenue to move beyond linear models without the loss of interpretable
statistical inference.
\end{abstract}

\section*{Keywords}
DNA methylation, differential methylation, epigenetics, causal inference, statistical variable importance, targeted minimum loss-based estimation, machine learning

\clearpage
\pagestyle{main}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

DNA methylation is a fundamental epigenetic process known to play an important
role in the regulation of gene expression. DNA CpG methylation involves the
addition of a methyl group (\(\text{CH}_3\)) to the fifth carbon of the cytosine
ring structure to form 5-methylcytosine. Numerous biological and medical studies
have implicated DNA CpG methylation as playing a role in disease and
development \citep{robertson2005dna}. Perhaps unsurprisingly then, biotechnologies
have been developed to study the molecular mechanisms of this epigenetic
process. Modern assays, like the Illumina \emph{Infinium} Methylation assay,
allow for quantitative interrogation of DNA methylation of CpG sites scattered
across the genome at single-nucleotide resolution; moreover, much effort
has been invested, by the bioinformatics community, in the development of tools
for properly removing technological effects that may contaminate biological
signatures measured by such assays. Despite these advances in both biological
and bioninformatical techniques, most statistical methods available for
differential analysis of data produced by such assays rely on over-simplified
(often generalized linear) models. This is because typical statistical
estimation approaches are not achievable for such high-dimensional data without
restrictive modeling assumptions. When these assumptions are violated, estimates
of the population-level effect of an exposure or treatment may have large bias.
Data-adaptive estimation procedures that utilize machine learning can control
for high-dimensional confounders when estimating treatment/exposure effects.
However, interpretable statistical inference (i.e.~confidence intervals,
hypothesis tests) resulting from these data-adaptive estimates is challenging
\citep{libbrecht2015machine}.

In this paper, we present an alternative such statistical analysis approaches in
the form of nonparametric estimation procedures that provide simple and well
interpretable statistical inference. Inspired by concepts of counterfactual
effects from the causal inference literature and machine learning algorithms, we
provide a computationally efficient technique for obtaining targeted estimates
of nonparametric \emph{variable importance measures} (\textbf{VIM}) for a pre-screened set
of CpGs, controlling for the False Discovery Rate (FDR) as if all sites were
tested. Under counterfactual assumptions, these estimators have asymptotic
normal sampling distributions and thus provide reliable inference.
Counterfactuals are ``counter''-fact, defined in terms of an unrealistic yet ideal
experiment. In the context of DNA methylation and for a binary treatment, we
define the counterfactual outcomes as the observed methylation value (whether
Beta- or M-values) a CpG site would have if all subjects were treated and the
methylation value a CpG site would have if all subjects were not treated. Of
course, these counterfactual outcomes are impossible to observe but they are
very helpful for us to asses the importance of a CpG site relative to a subjects
treatment/exposure and they can be statistically estimated
\citep{vdl2006statistical}. Our VIM definition uses the concepts of counterfactuals
and the framework for estimation is built on \emph{targeted maximum likelihood
estimation} or \emph{targeted minimum loss-based estimation} (\textbf{TMLE})
\citep{vdl2006targeted}. This methodology targets the individual importance of each
CpG site by taking advantage of the prediction power of machine learning
algorithms. The mathematical properties of this approach have been explored in
\citep{tuglus2008targeted} and we focus on our contribution to it's application,
\texttt{methyvim}, an R package that make this technique for differential methylation
analysis easily accessible.

For a general discussion of the framework of targeted minimum loss-based
estimation and the role this approach plays in statistical causal inference, the
interested reader is invited to consult \citet{vdl2011targeted} and \citet{vdl2017targeted}.
For a more general introduction to (statistical) causal inference,
\citet{pearl2009causality} and \citet{hernan2018causal} may be of interest.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

The core functionality of this package is made available via the eponymous
\texttt{methyvim} function, which implements a statistical algorithm designed to
compute targeted estimates of VIMs, defined in such a way that the VIMs
represent parameters of scientific interest in computational biology
experiments; moreover, these VIMs are defined such that they may be estimated in
a manner that is very nearly assumption-free, that is, within a \emph{fully
nonparametric statistical model}. \textbf{The statistical algorithm consists in
several major steps}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Pre-screening of genomic sites is used to isolate a subset of sites for
  which there is cursory evidence of differential methylation. For the sake of
  computational feasibility, targeted minimum loss-based estimates of VIMs are
  computed only for this subset of sites. Currently, the available screening
  approach adapts core routines from the
  \href{http://bioconductor.org/packages/limma}{\texttt{limma}} R
  package. Future releases will support functionality from other packages
  (e.g., \href{https://CRAN.R-project.org/package=randomForest}{\texttt{randomForest}},
  \href{https://CRAN.R-project.org/package=tmle.npvi}{\texttt{tmle.npvi}}). Following the
  style of the function for performing screening via \texttt{limma}, users may write
  their own screening functions and are invited to contribute such functions to
  the core software package by opening pull requests at the GitHub repository.
\item
  Nonparametric estimates of VIMs, for the specified target parameter, are
  computed at each of the CpG sites passing the screening step. The VIMs are
  defined in such a way that the estimated effects is of an exposure/treatment
  on the methylation status of a target CpG site, controlling for the observed
  methylation status of the neighbors of that site. Currently, routines are
  adapted from the \href{https://CRAN.R-project.org/package=tmle}{\texttt{tmle}} R package.
  Future releases will support doubly-robust estimates of these VIMs (via the
  \href{https://cran.r-project.org/package=drtmle}{\texttt{drtmle}} package)
  and add parameters for continuous treatments/exposures (via the
  \href{https://CRAN.R-project.org/package=tmle.npvi}{\texttt{tmle.npvi}} package).
\item
  Since pre-screening is performed prior to estimating VIMs, we make use of a
  multiple testing correction uniquely suited to such settings. Due to the
  multiple testing nature of the estimation problem, a variant of the Benjamini
  \& Hochberg procedure for controlling the False Discovery Rate (FDR) is
  applied \citep{benjamini1995controlling}. Specifically, we apply the ``modified
  marginal Benjamini \& Hochberg step-up False Discovery Rate controlling
  procedure for multi-stage analyses'' (FDR-MSA), which is guaranteed to
  control the FDR as if all sites were tested (i.e., without screening)
  \citep{tuglus2009modified}.
\end{enumerate}

\hypertarget{parameters-of-interest}{%
\paragraph{Parameters of Interest}\label{parameters-of-interest}}

For the CpG sites that passed the pre-screening step, the user chooses the
parameter of interest that will be computed at each site. \emph{In all cases, an
estimator of the parameter of interest is constructed via targeted minimum
loss-based estimation}.

For discrete-valued treatments or exposures:

\begin{itemize}
\item
  The average treatment effect (ATE): The effect of a binary exposure or
  treatment on the observed methylation at a target CpG site is estimated,
  controlling for the observed methylation at all other CpG sites in the same
  neighborhood as the target site, based on an additive form. Often denoted
  \(\psi_0 = \psi_0(1) - \psi_0(0)\), the parameter estimate represents the
  additive difference in methylation that would have been observed at the
  target site had all observations received the treatment versus the
  counterfactual under which none received the treatment.
\item
  The relative risk (RR): The effect of a binary exposure or treatment on the
  observed methylation at a target CpG site is estimated, controlling for the
  observed methylation at all other CpG sites in the same neighborhood as the
  target site, based on a geometric form. Often denoted,
  \(\psi_0 = \frac{\psi_0(1)}{\psi_0(0)}\), the parameter estimate represents the
  multiplicative difference in methylation that would have
  been observed at the target site had all observations received the treatment
  versus the counterfactual under which none received the treatment.
\end{itemize}

Estimating the VIM corresponding to the parameters above, for discrete-valued
treatments or exposures, requires two separate regression steps: one for the
treatment mechanism (propensity score) and one for the outcome regression.
Technical details on the nature of these regressions are discussed in
\citet{hernan2018causal}, and details for estimating these regressions in the
framework of targeted minimum loss-based estimation are discussed in
\citet{vdl2011targeted}.

Support for continuous-valued treatments or exposures is \emph{planned but not yet
available}, though work is underway to incorporate into our methodology the
following

\begin{itemize}
\item
  A nonparametric variable importance measure (NPVI) \citep{chambaz2012estimation}:
  The effect of continuous-valued exposure or treatment (the observed
  methylation at a target CpG site) on an outcome of interest is estimated,
  controlling for the observed methylation at all other CpG sites in the same
  neighborhood as the target (treatment) site, based on a parameter that
  compares values of the treatment against a reference value taken to be the
  null. In particular, the implementation provided is designed to assess the
  effect of differential methylation at the target CpG site on a (typically)
  phenotype-level outcome of interest (e.g., survival), in effect providing an
  nonparametric evaluation of the impact of methylation at the target site on
  said outcome.
\end{itemize}

\hypertarget{class-methytmle}{%
\paragraph{\texorpdfstring{Class \texttt{methytmle}}{Class methytmle}}\label{class-methytmle}}

We have adopted a class \texttt{methytmle} to help organize the functionality within
this package. The \texttt{methytmle} class builds upon the \texttt{GenomicRatioSet} class
provided by the \texttt{minfi} package so all of the slots of \texttt{GenomicRatioSet} are
contained in a \texttt{methytmle} object. The new class introduced in the \texttt{methyvim}
package includes several new slots:

\begin{itemize}
\item
  \texttt{call} - the form of the original call to the \texttt{methyvim} function.
\item
  \texttt{screen\_ind} - indices identifying CpG sites that pass the screening process.
\item
  \texttt{clusters} - non-unique IDs corresponding to the manner in wich sites are
  treated as neighbors. These are assigned by genomic distance (bp) and respect
  chromosome boundaries (produced via a call to \texttt{bumphunter::clusterMaker}).
\item
  \texttt{var\_int} - the treatment/exposure status for each subject. Currently, these
  must be binary, due to the definition of the supported targeted parameters.
\item
  \texttt{param} - the name of the target parameter from which the estimated VIMs are
  defined.
\item
  \texttt{vim} - a table of statistical results obtained from estimating VIMs for
  each of the CpG sites that pass the screening procedure.
\item
  \texttt{ic} - the measured array values for each of the CpG sites passing the
  screening, transformed into influence curve space based on the chosen target
  parameter.
\end{itemize}

We refer the reader to the package vignette, ``\texttt{methyvim}: Targeted Data-Adaptive
Estimation and Inference for Differential Methylation Analysis,'' included in any
distribution of the software package, for further details.

\hypertarget{operation}{%
\subsection{Operation}\label{operation}}

A standard computer with the latest version of R and Bioconductor 3.6 installed
will handle applications of the \texttt{methyvim} package.

\hypertarget{use-cases}{%
\section{Use Cases}\label{use-cases}}

To examine the practical applications and the full set of utilities of the
\texttt{methyvim} package, we will use a publicly available example data set produced
by the Illumina 450K array, from the \texttt{minfiData} R package.

\hypertarget{preliminaries-setting-up-the-data}{%
\paragraph{Preliminaries: Setting up the Data}\label{preliminaries-setting-up-the-data}}

We begin by loading the package and the data set. After loading the data, which
comes in the form of a raw \texttt{MethylSet} object, we perform some further
processing by mapping to the genome (with \texttt{mapToGenome}) and converting the
values from the methylated and unmethylated channels to Beta-values
(via \texttt{ratioConvert}). These two steps together produce an object of class
\texttt{GenomicRatioSet}, provided by the \texttt{minfi} package.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{suppressMessages}\NormalTok{(}\KeywordTok{library}\NormalTok{(minfiData))}
\KeywordTok{data}\NormalTok{(MsetEx)}
\NormalTok{mset <-}\StringTok{ }\KeywordTok{mapToGenome}\NormalTok{(MsetEx)}
\NormalTok{grs <-}\StringTok{ }\KeywordTok{ratioConvert}\NormalTok{(mset)}
\NormalTok{grs}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## class: GenomicRatioSet 
## dim: 485512 6 
## metadata(0):
## assays(2): Beta CN
## rownames(485512): cg13869341 cg14008030 ... cg08265308 cg14273923
## rowData names(0):
## colnames(6): 5723646052_R02C02 5723646052_R04C01 ...
##   5723646053_R05C02 5723646053_R06C02
## colData names(13): Sample_Name Sample_Well ... Basename filenames
## Annotation
##   array: IlluminaHumanMethylation450k
##   annotation: ilmn12.hg19
## Preprocessing
##   Method: Raw (no normalization or bg correction)
##   minfi version: 1.21.2
##   Manifest version: 0.4.0
\end{verbatim}

We can create an object of class \texttt{methytmle} from any \texttt{GenomicRatioSet} object
simply invoking the S4 class constructor:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(methyvim)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## methyvim v1.1.0: Targeted Variable Importance for Differential Methylation Analysis
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mtmle <-}\StringTok{ }\KeywordTok{.methytmle}\NormalTok{(grs)}
\end{Highlighting}
\end{Shaded}

Additionally, a \texttt{GenomicRatioSet} can be created from a matrix with the
function \texttt{makeGenomicRatioSetFromMatrix} provided by the \texttt{minfi} package.

\hypertarget{differential-methylation-analysis}{%
\paragraph{Differential Methylation Analysis}\label{differential-methylation-analysis}}

For this example analysis, we'll treat the condition of the patients as the
exposure/treatment variable of interest. The \texttt{methyvim} function requires that
this variable either be \texttt{numeric} or easily coercible to \texttt{numeric}. To
facilitate this, we'll simply convert the covariate (currently a \texttt{character}):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{var_int <-}\StringTok{ }\NormalTok{(}\KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{as.factor}\NormalTok{(}\KeywordTok{colData}\NormalTok{(grs)}\OperatorTok{$}\NormalTok{status)) }\OperatorTok{-}\StringTok{ }\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\textbf{n.b.}, the re-coding process results in ``normal'' patients being assigned a
value of 1 and cancer patients a 0.

Now, we are ready to analyze the effects of cancer status on DNA methylation
using this data set. We proceed as follows with a targeted minimum loss-based
estimate of the Average Treatment Effect.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{suppressMessages}\NormalTok{(}
\NormalTok{  methyvim_cancer_ate <-}\StringTok{ }\KeywordTok{methyvim}\NormalTok{(}\DataTypeTok{data_grs =}\NormalTok{ grs, }\DataTypeTok{var_int =}\NormalTok{ var_int,}
                                  \DataTypeTok{vim =} \StringTok{"ate"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"Beta"}\NormalTok{, }\DataTypeTok{filter =} \StringTok{"limma"}\NormalTok{,}
                                  \DataTypeTok{filter_cutoff =} \FloatTok{0.20}\NormalTok{, }\DataTypeTok{obs_per_covar =} \DecValTok{2}\NormalTok{,}
                                  \DataTypeTok{parallel =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{sites_comp =} \DecValTok{125}\NormalTok{,}
                                  \DataTypeTok{tmle_type =} \StringTok{"glm"}
\NormalTok{                                 )}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in set_parallel(parallel = parallel, future_param = future_param, : Sequential evaluation is strongly discouraged. 
##  Proceed with caution.
\end{verbatim}

Note that we set the \texttt{obs\_per\_covar} argument to a relatively low value (2,
where the recommended default is 20) for the purposes of this example as the
sample size is only 10. We do this only to exemplify the estimation procedure
and it is important to point out that such low values for \texttt{obs\_per\_covar} will
compromise the quality of inference obtained because this setting directly
affects the definition of the target parameter.

Further, note that here we apply the \texttt{glm} flavor of the \texttt{tmle\_type} argument,
which produces faster results by fitting models for the propensity score and
outcome regressions using a limited number of parametric models. By contrast,
the \texttt{sl} (for ``Super Learning'') flavor fits these two regressions using highly
nonparametric and data-adaptive procedures (i.e., via machine learning).
Obtaining the estimates via GLMs results in each of the regression steps
being less robust than if nonparametric regressions were used.

We can view a table of results by examining the \texttt{vim} slot of the produced
\texttt{methytmle} object.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(}\KeywordTok{slot}\NormalTok{(methyvim_cancer_ate, }\StringTok{"vim"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                 lwr_ci      est_ate       upr_ci      var_ate         pval
## cg14008030 -0.11956597 -0.031415962  0.056734049 2.022705e-03 4.848468e-01
## cg20253340 -0.08850637 -0.058866142 -0.029225917 2.286919e-04 9.917426e-05
## cg21870274 -0.09499057 -0.029118982  0.036752609 1.129495e-03 3.862537e-01
## cg17308840 -0.04626018 -0.007152452  0.031955277 3.981191e-04 7.199943e-01
## cg00645010 -0.02697515 -0.013809263 -0.000643373 4.512199e-05 3.980386e-02
## cg27534567  0.06745648  0.115711854  0.163967223 6.061486e-04 2.602936e-06
##            n_neighbors n_neighbors_control max_cor_neighbors
## cg14008030           0                   0                NA
## cg20253340           0                   0                NA
## cg21870274           2                   1         0.9443580
## cg17308840           2                   1         0.9443580
## cg00645010           2                   2         0.5236810
## cg27534567           1                   0         0.9362968
\end{verbatim}

Finally, we may compute FDR-corrected p-values, by applying a modified procedure
for controlling the False Discovery Rate for multi-stage analyses (FDR-MSA)
\citep{tuglus2009modified}. We do this by simply applying the \texttt{fdr\_msa} function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fdr_p <-}\StringTok{ }\KeywordTok{fdr_msa}\NormalTok{(}\DataTypeTok{pvals =} \KeywordTok{slot}\NormalTok{(methyvim_cancer_ate, }\StringTok{"vim"}\NormalTok{)}\OperatorTok{$}\NormalTok{pval,}
                 \DataTypeTok{total_obs =} \KeywordTok{nrow}\NormalTok{(methyvim_cancer_ate))}
\end{Highlighting}
\end{Shaded}

Having explored the results of our analysis numerically, we now proceed to use
the visualization tools provided with the \texttt{methyvim} R package to further
enhance our understanding of the results.

\hypertarget{visualization-of-results}{%
\paragraph{Visualization of Results}\label{visualization-of-results}}

While making allowance for users to explore the full set of results produced by
the estimation procedure (by way of exposing these directly to the user), the
\texttt{methyvim} package also provides \emph{three} (3) visualization utilities that
produce plots commonly used in examining the results of differential methylation
analyses.

A simple call to \texttt{plot} produces side-by-side histograms of the raw p-values
computed as part of the estimation process and the corrected p-values obtained
from using the FDR-MSA procedure.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(methyvim_cancer_ate, }\DataTypeTok{type =} \StringTok{"raw_pvals"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{paper_BiocF1000_files/figure-latex/methyvim-pvals-raw-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(methyvim_cancer_ate, }\DataTypeTok{type =} \StringTok{"fdr_pvals"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{paper_BiocF1000_files/figure-latex/methyvim-pvals-fdr-1} \end{center}

\textbf{Remark:} The plots displayed above may also be generated as side-by-side
histograms in a single plot object. This is the default for the \texttt{plot} method
and may easily be invoked by specifying no additional arguments to the \texttt{plot}
function, unlike in the above.

While histograms of the p-values may be generally useful in inspecting the
results of the estimation procedure, a more common plot used in examining the
results of differential methylation procedures is the volcano plot, which plots
the parameter estimate along the x-axis and \(-\text{log}_{10}(\text{p-value})\)
along the y-axis. We implement such a plot in the \texttt{methyvolc} function:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{methyvolc}\NormalTok{(methyvim_cancer_ate)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{paper_BiocF1000_files/figure-latex/methyvim-volcano-1} \end{center}

The purpose of such a plot is to ensure that very low (possibly statistically
significant) p-values do not arise from cases of low variance. This appears to
be the case in the plot above (notice that most parameter estimates are near
zero, even in cases where the raw p-values are quite low).

Yet another popular plot for visualizing effects in such settings is the
heatmap, which plots estimates of the raw methylation effects (as measured by
the assay) across subjects using a heat gradient. We implement this in the
\texttt{methyheat} function:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{methyheat}\NormalTok{(methyvim_cancer_ate)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{paper_BiocF1000_files/figure-latex/methyvim-heatmap-1} \end{center}

Invoking \texttt{methyheat} in this manner produces a plot of the top sites (\(25\), by
default) based on the raw p-value, using the raw methylation measures in the
plot. This uses the exceptional \texttt{superheat} R package \citep{barter2017superheat}.

\hypertarget{summary}{%
\section{Summary}\label{summary}}

Here we introduce the R package \texttt{methyvim}, an algorithm for differential
methylation analysis that moves beyond linear models. Additionally, this
estimation procedure produces straightforward statistical inference. Inspired by
concepts of counterfactual effects from the causal inference literature and
machine learning algorithms, we provide a computationally efficient technique
for obtaining targeted estimates of nonparametric variable importance measures
for a pre-screened set of CpGs, controlling for the False Discovery Rate as if
all sites were tested, and the estimation framework is built on targeted
minimum loss-based estimation. By taking advantage of the prediction power of
machine learning algorithms, this methodology targets the individual importance
of each CpG site, providing a fully nonparametric statistical model for the
analysis of differential methylation.

\hypertarget{software-availability}{%
\section{Software availability}\label{software-availability}}

Latest source code (development version): \url{https://github.com/nhejazi/methyvim}

Bioconductor (stable release): \url{https://bioconductor.org/packages/methyvim}

Archived source code as at time of
publication: \url{https://github.com/nhejazi/methyvim}

Software license: The MIT License, copyright Nima S. Hejazi

\hypertarget{author-contributions}{%
\section{Author contributions}\label{author-contributions}}

NH designed and implemented the software package, applied the tool to the use
cases presented, and co-drafted the present manuscript. RP helped in designing
the software and co-drafted the present article. AH and ML served as advisors
for the development of this software and general statistical algorithm.

\hypertarget{competing-interests}{%
\section{Competing interests}\label{competing-interests}}

No competing interests were disclosed.

\hypertarget{grant-information}{%
\section{Grant information}\label{grant-information}}

NH was supported in part by the National Library of Medicine of the National
Institutes of Health under Award Number T32-LM012417, by P42-ES004705, and
by R01-ES021369. RP was supported by P42-ES004705. The content of this work is
solely the responsibility of the authors and does not necessarily represent the
official views of the various funding sources and agencies.

\renewcommand\refname{References}
{\small\bibliography{paper\_BiocF1000.bib}}

\end{document}